---
title: "DATA607 - Assignment1"
author: "Michael Ippolito"
date: "2/4/2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview  
  
In their article, Why Americans Don't Vote, Thomson-Deveaux et al. (2020) explored the reasons why a large number of eligible voters (35 to 60 percent) don't vote in US elections and at the voting habits of voters broken out by various categories (age, level of education, race, gender, and income). The data collected confirmed the well-accepted notion that older, more educated people with higher incomes and stronger party affiliations are more likely to vote. Of voters who never, rarely, or only sometimes vote in US elections, Thomson-Deveaux et al. reported the top reasons why that was the case. My assignment will focus on the specific polling question that dealt with the reason why people often don't vote, categorized by the frequency with which they do vote.  
  
Article citation:  
FiveThirtyEight (2020). Why Americans Don't Vote. https://projects.fivethirtyeight.com/non-voters-poll-2020-election/.  
  
## Load libraries, retrieve data from github, and parse CSV  

```{r warning=FALSE}

library(tidyverse)
library(RCurl)
#original URL on fivethirtyeight,com:
#https://raw.githubusercontent.com/fivethirtyeight/data/master/non-voters/nonvoters_data.csv
nonvoters_csv <- getURL("https://raw.githubusercontent.com/mmippolito/cuny/main/data607/assignment1/nonvoters_data.csv")
nonvoters <- read.csv(text = nonvoters_csv)

```

## Revelant variables  
  
The data included the responses from all survey questions, tabulated in 119 variables and 5,836 observations. For this assignment, the subset I chose included the following variables:  
  
| voter_category        How often the voter voted:  
|                         always  
|                         sporadic  
|                         rarely/never  
|   
| Q29                   Survey question #29, which asked voters to mark which of the following   
|                       ten reason were most important factors in why they chose not to vote.
|                       (A value of 1 indicates the voter marked it as important; a value of -1 means
|                       they answered the question, but didn't mark this answer as important; and N/A
|                       means they didn't answer the question.)
|  
|                         Q29_1   I didn't like any of the candidates  
|                         Q29_2   Because of where I live, my vote doesn't matter  
|                         Q29_3   No matter who wins, nothing will change for people like me  
|                         Q29_4   Our system is too broken to be fixed by voting  
|                         Q29_5   I wanted to vote, but I didn't have time, couldn't get off work,  
|                                 something came up, or I forgot  
|                         Q29_6   I'm not sure if I can vote  
|                         Q29_7   Nobody talks about the issues that are important to me personally  
|                         Q29_8   All the candidates are the same  
|                         Q29_9   I don't believe in voting  
|                         Q29_10  Other  
  
## Create Q29 array
  
```{r}

# Make Q29 into an array so it can be later iterated over in a "for" loop
q29 <- c(
  "I didn't like any of the candidates",
  "Because of where I live, my vote doesn't matter",
  "No matter who wins, nothing will change for people like me",
  "Our system is too broken to be fixed by voting",
  "I wanted to vote, but I didn't have time, couldn't get off work, something came up, or I forgot",
  "I'm not sure if I can vote",
  "Nobody talks about the issues that are important to me personally",
  "All the candidates are the same",
  "I don't believe in voting",
  "Other"
)

```

## Create subset from relevant variables  
  
First, filter out voters who "always" vote, and only select voter_category, weight, and the Question 29 responses.  

```{r result='asis'}

# Filter out voters who always vote; only select specific variables
nonvoters_29 <- as_tibble(select(nonvoters, voter_category, weight, Q29_1:Q29_10)) %>% 
  filter(voter_category != "always")
nonvoters_29[1:5,]     # Display first 5 observations

```

## Pivot on Q29 responses

Now create 10 different tibbles--one for each answer in question 29--and group on voter category.

```{r echo=TRUE, results='asis'}

# Iterate over each answer in Question 29
for(i in 1:10) {

  # Concatenate i to create variable name string
  q <- paste("Q29_", i, sep = "")

  # Create tibble with weighted count of voters who answered that this was
  # an important reason why they didn't vote
  categories <- select(nonvoters_29, voter_category, weight, i + 2) %>% 
    filter(!is.na(get(q)) & get(q) == 1) %>%
    group_by(voter_category) %>%
    summarize(wt = sum(weight))

  # Create tibble with weighted counts of voters who answered this question at all
  totals <- select(nonvoters_29, voter_category, weight, i + 2) %>% 
    filter(!is.na(get(q))) %>%
    group_by(voter_category) %>%
    summarize(wt_total = sum(weight))

  # Merge the two tibbles
  subset <- merge(categories, totals, by = "voter_category")
  
  # Create a new variable for percentage and print the new tibble
  subset <- mutate(subset, percentage = wt * 100 / wt_total)
  print(subset)

  # Plot the bar chart
  print(ggplot(data = subset, mapping = aes(x = voter_category, y = percentage)) + 
    geom_bar(stat = "identity", mapping = aes(color = voter_category, fill = voter_category)) + 
    ggtitle(q29[i]) + 
    theme(plot.title = element_text(hjust = 0.5)))

}

```

## Conclusions
  
As evidenced by the data, voters who never, rarely, or only sporadically vote tend to feel as if nothing will change, regardless of the outcome of an election. Almost as often, they report not feeling any affinity toward any particular candidate. Further, many of them claim that the system is in disrepair and can't be fixed by voting.
  
While the above results are interesting (albeit disheartening!), I'd find it even more telling to further break down the most significant response by gender, race, income, or level of education; this might indicate which voters feel disenfranchised and why they feel that way, thereby guiding public policy decisions on possibly mitigation efforts.

## People who feel as if nothing will change  
  
Break the results of response #3 of question 29 out by age, education, race, gender, and income.

```{r result='asis'}

# Filter out voters who always vote; only select response #3 (people who feel as if nothing will change)
nonvoters_29_3 <- 
  as_tibble(select(nonvoters, voter_category, weight, ppage, educ, race, gender, income_cat, Q29_3)) %>% 
  filter(voter_category != "always" & !is.na(Q29_3))

#cut the age variable into categories
nonvoters_29_3 <- mutate(nonvoters_29_3, age_category = cut(ppage, c(18, 30, 40, 50, 60, 70, 80, 120)))
nonvoters_29_3[1:5,]     # Display first 5 observations

# Create array of fields we're interestedin
fields <- c("age_category", "educ", "race", "gender", "income_cat")

# Iterate over each answer in Question 29
for(i in 1:5) {

  # Create tibble with total weighted counts of voters
  totals <- select(nonvoters_29_3, weight, fields[i]) %>% 
    filter(!is.na(fields[i])) %>%
    group_by_at(fields[i]) %>%
    summarize(wt_total = sum(weight))

  # Create tibble with weighted count of voters who answered that this was
  # an important reason why they didn't vote
  grouped <- select(nonvoters_29_3, weight, Q29_3, fields[i]) %>% 
    filter(!is.na(fields[i]) & Q29_3 == 1) %>%
    group_by_at(fields[i]) %>%
    summarize(wt = sum(weight))

  # Merge the two tibbles
  subset <- merge(grouped, totals, by = fields[i])

  # Create a new variable for percentage and print the new tibble
  subset <- mutate(subset, percentage = wt * 100 / wt_total)
  print(subset)

  # Plot the bar chart
  print(ggplot(data = subset, mapping = aes(x = fields[i], y = percentage)) + 
    geom_bar(position = "dodge", stat = "identity", mapping = aes(color = get(fields[i]), fill = get(fields[i]))) + 
    ggtitle(fields[i]) + 
    theme(plot.title = element_text(hjust = 0.5)))

}

```

## Further Conclusions
  
The additional analysis for Response #3 illustrate which demographics feel their vote doesn't matter. Generally speaking, there was little variation across demographics, with one notable exception: While older voters (70 and older) tend to vote more often than younger ones, they are also the ones who feel their vote doesn't matter. While this inverse relationship is somewhat surprising, the fact that there was little variation among the other demographics is perhaps even more surprising. For example, my expectation was that minorities and females would comprise a greater percentage of voters who feel disengranchised, given the demographic of most elected politicians. Instead, there was little variation among races and genders, which indicates to me that people, in general, feel their vote doesn't matter, regardless of demographic. In a time when there seems to be little to unify us, perhaps we can console ourselves by being united by our disillusionment.
