---
title: "DATA607 - Assignment 3"
author: "Michael Ippolito"
date: "2/19/2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
# Init
knitr::opts_chunk$set(echo = TRUE)
options(dplyr.summarise.inform = FALSE)
library(tidyverse)
library(knitr)
library(RCurl)
```
  
**1. Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"**  
  
```{r problem_1}
#original URL on fivethirtyeight,com:
#https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv
majors_csv <- getURL("https://raw.githubusercontent.com/mmippolito/cuny/main/data607/assignment3/majors-list.csv")
majors <- read.csv(text = majors_csv)
majors_subset <- str_subset(majors$Major, "(DATA|STATISTICS)")
majors_subset
```
  
  
**2 Write code that transforms the data below:**  
  
[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
  
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
  
[13] "olive"        "salal berry"
  
Into a format like this:
  
c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")
  
```{r problem_2}
raw_string <- '
[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
  
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
  
[13] "olive"        "salal berry"
'
matches <- str_match_all(raw_string, '"(.+?)"\\s+')
vectored_list <- matches[[1]][,2]
vectored_list
```
  
  
**3 Describe, in words, what these expressions will match:**  
  
* (.)\\1\\1  
  
Any three characters in a row, e.g. AA  

```{r problem3_1}

s <- "abcAAAabc"
str_extract(s, "(.)\\1\\1")

```
  
* "(.)(.)\\\\2\\\\1"  
  
A four-character palindrome, e.g ABBA  

```{r problem3_2}

s <- "abcABBAabc"
str_extract(s, "(.)(.)\\2\\1")

```
  
* (..)\\1  
  
A set of two repeating characters, e.g. ABAB  

```{r problem3_3}

s <- "abcABABabc"
str_extract(s, "(..)\\1")

```

* "(.).\\\\1.\\\\1"  
  
The same character repeated three times, separated by any other character in between, e.g. AxAyA  
  
```{r problem3_4}

s <- "abcAxAyAabc"
str_extract(s, "(.).\\1.\\1")

```

* "(.)(.)(.).*\\\\3\\\\2\\\\1"  
  
A six-character palindrome separated in the middle by any number of characters, e.g. ABCxyzCBA  
  
```{r problem3_5}

s <- "abcABCxyzCBAabc"
str_extract(s, "(.)(.)(.).*\\3\\2\\1")

```

  
**4 Construct regular expressions to match words that:**  
  
* Start and end with the same character.  

^(.).*\\1$  

```{r problem4_1}

s <- "AasdfasdfA"
str_extract(s, "^(.).*\\1$")

```

* Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)  

([A-Za-z]{2}).*\\1  

```{r problem4_2}

s <- "church"
str_extract(s, "([A-Za-z]{2}).*\\1")

```

* Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)  

([A-Za-z]).*\\1.*\\1  
  
```{r problem4_3}

s <- "eleven"
str_extract(s, "([A-Za-z]).*\\1.*\\1")

```
