---
title: "DATA607 - Assignment 9"
author: "Michael Ippolito"
date: "4/11/2021"
output:
  html_document:
    df_print: paged
    toc: false
    toc_float: true
  pdf_document: default
---

```{r setup, include=FALSE}

# Init
knitr::opts_chunk$set(echo = TRUE)
options(dplyr.summarise.inform = FALSE)
library(tidyverse)
library(knitr)
library(RCurl)
library(kableExtra)
library(RJSONIO)
library(getPass)

```

## Assignment 7 {.tabset}

### Overview

The assignment was to do the following:

1. Sign up for an API key on the New York Times website.
2. Build a procedure in R to read from the API.
3. Store the data into an R data frame.


### Approach

My approach is follows:

* Sign up for an API key, register a new application, and enable the Top Stories API.
* Fetch the "world" top stories from the site using the API, storing it locally.
* Read and parse the JSON file, tidying the data as necessary.
* Store and display the data in a data frame.

### Fetch data

```{r fetch data}

if (interactive()) {
  api_key <- getPass("Enter API key:")
} else {
  api_key = readChar('api_key.txt', file.info('api_key.txt')$size)
}

# To prevent reading from the NYT site continuously, first see if the data has already been read
if (!file.exists("nyt_world.json")) {
  # Fetch JSON from NYT site
  raw_json <- getURL(str_c("https://api.nytimes.com/svc/topstories/v2/world.json?api-key=", api_key))
  write(raw_json, "nyt_world.json")
}

# Read from saved JSON
rawjson <- fromJSON("nyt_world.json", simplifyVector = TRUE)

# Extract column names
vars <- names(rawjson[['results']][[1]])

```

### Main data frame

```{r main_df}

# Create main data frame
df <- data.frame(matrix(ncol = length(vars), nrow = length(rawjson[['results']])))
colnames(df) <- vars

# Remove multimedia column; a separate child table will be created for this
df <- df %>% select(-multimedia)

# Create the id field as the primary key that will link the multimedia table
df <- df %>% mutate(id = c(1:length(rawjson[['results']])))

# Walk each variable, assigning each one to a column of the empty data frame
for (v in vars) {

  # Extract this column to a temporary variable
  tmpelement <- sapply(rawjson[['results']], '[[', v)

  # Handle different type of elements
  if (v == "multimedia") {
    # Handle mutlimedia separately
  }
  else if (typeof(tmpelement) == "list") {
    #If this element is a list, collapse it into a single value
    df[v] <- sapply(tmpelement, paste, collapse = "|")
  }
  else {
    df[v] <- tmpelement
  }

}

# Replace UTF-8 quote marks
df$title <- str_replace_all(df$title, "’", "'")
df$abstract <- str_replace_all(df$title, "’", "'")
df$title <- str_replace_all(df$title, "‘", "'")
df$abstract <- str_replace_all(df$title, "‘", "'")

```

### Multimedia data frame

```{r mm_df}

# Create mutlimedia data frame
mmvars <- names(rawjson[['results']][[1]]['multimedia'][[1]][[1]])
mmvars <- c('id', mmvars)
df_mm = data.frame(matrix(vector(), 0, length(mmvars), dimnames = list(c(), mmvars)), stringsAsFactors=F)

# Multimedia
for (i in 1:length(rawjson[['results']])) {
  for (j in 1:length(rawjson[['results']][[i]]['multimedia'][[1]])) {
    dftmp = data.frame(rawjson[['results']][[i]]['multimedia'][[1]][j])
    dftmp <- dftmp %>% mutate(id = i)
    df_mm <- rbind(df_mm, dftmp)
  }
}
    
```

### Display data frames

```{r dataframes}

# Display dataframes
kable(df, caption = 
  "<i><font color=#000000><b>Table 1.</b> Main data frame</font></i>") %>% 
  kable_styling(latex_options = "striped")
kable(df_mm, caption = 
  "<i><font color=#000000><b>Table 2.</b> Multimedia data frame</font></i>") %>% 
  kable_styling(latex_options = "striped")

```

### Conclusion

This was a relatively straightforward process, except that manipulating lists of lists in R continues to confound me. There is probably a way of using lapply or sapply instead of the nested for loops above, but I gave up after about 4 hours of trying. The other small problem was that the unicode left-quote and right-quote characters appeared in some of the fields and had to be converted. An additional consideration was how to handle some of the data that arrived as lists of elements. I chose to create a child table for the multimedia data. But for the other data, I concatenated them into a single field, using pipes as separators.
